Parsing Tweets into Universal Dependencies

### 2017-07-22 paper.tex
lp pos da 2

\subsection{Linguistic Phenomena of Twitter} \label{introduction}
%Eisenstan
As the representative of the web language, different from the standard languages such as newswire, twitter has its own linguistic phenomena, that could be categorized into word level and structure level phenomena, and we will discuss them separately.

@@

There are special tokens that will appear only or much more frequently in tweets. According to the syntactic functions these tokens have in the tweets, we classify them into three types: syntactic tokens, non-syntactic tokens and optionally syntactic tokens. 
Syntactic tokens are tokens that will always have syntactic functions in tweets, but they might not be in their normal orthographic form and we need to do some preprocessing to recover their standard form. Then we can treat them just as normal tokens in the standard text.
Non-syntactic tokens are tokens that will not have syntactic function or semantic meaning in tweets, and they are not considered in the syntax analysis and need to be treated independently.
Optionally syntactic tokens are tokens that can have either syntactic or non-syntactic functions based on the context. Usually they are tokens like URLs ad hashtags that do not participate in the syntax of the tweets, and we treat them as non-syntactic functions by default. However, similar in the analysis of \newcite{kong-EtAl:2014:EMNLP2014}, in some cases they are perceived as the arguments or predicates of tweets and therefore have syntactic functions. In this case we treat them as syntactic tokens.

@@

\begin{itemize}
\item syntactic tokens
	\begin{itemize}
		\item phrasal abbreviations: {\tt wtf} (what the fuck), {\tt rn} (right now), etc.
		\item orthographic variants: {\tt u} (you), {\tt sooooo}, {\tt dat} (that), etc.
	\end{itemize}
\item Non-syntactic tokens
	\begin{itemize}
		\item truncated tokens: last tokens that are truncated in tweets exceeding twitter character limits. It might be possible to recover the original tokens from the context in some cases, it is still extremely hard to further predict the rest of the truncated part.
	\end{itemize}
\item Optionally syntactic tokens
	\begin{itemize}
		\item retweet discourse marker: RT
		\item URL: http://bit.ly/xyz
		\item hashtag: \#ACL
		\item at-mentioned username: @user
		\item emoticons and emojis: :), :-), $\heart$, etc.
	\end{itemize}
\end{itemize}

As optionally syntactic tokens will be treated as syntactic or non-syntactic tokens based on the specific context, for the sake of simplicity, we will only discuss the syntactic and non-syntactic tokens in the following sections.

@@

\begin{itemize}
\item Retweet structure: {\tt RT @user : \{tweet content\}} is a typical structural pattern when a user is retweeting from other users. 
\item At-mentioned structure: Except the retweet structure and syntactically at-mentioned username, we think that all of at-mentioned usernames are representing a vocative situation, usually appearing at the beginning or the end of the sentences in tweets.
\item Paratactic parts without punctuations: We use ``part'' to represent a self-contained clause or phrase, and is independent of the previous and the following part. Very often, one tweet is comprised of multiple parts without any delimiting punctuations, such as {\tt \{part 1\} \{part 2\} ...}
\end{itemize}

We treat retweet structure and at-mention structure as structural patterns and keep their annotations consistent across tweets. For parataxis case, we treat each single sentence or phrases independently, and connect them together following the UD conventions.

Consider the following example

@@

From the example, we argue that {\tt u} and {\tt rn} are syntactic tokens.
{\tt th} is a truncated token and does not have any syntactic function, although we can infer that the {\tt th} could be probably {\tt the}.
{\tt \#ACL2017}, {\tt :)} and {\tt http://url} are optionally syntactic tokens, and should be all treated as non-syntactic tokens in this example.
{\tt RT @Yijia :} should be analyzed as the retweet structure, and {\tt heading to Canada for ACL} and {\tt can u gimme some money} should be considered as paratactic clauses without punctuations.

@@

\subsection{POS Tagging}
We follow the UD V2 morphology guideline and use the Universal POS tags~\cite{PETROV12.274} to tag the tweet tokens. For twitter specific linguistic phenomena, we have set different strategies and try to align with the UD conventions and UD\_English as close as possible.\footnote{https://github.com/UniversalDependencies/UD\_English}

@@

In structure level phenomena, we think {\tt RT @user} in retweet structure have no contribution to the tweet syntax and tag both of them as X.
We tag the at-mentioned username in at-mentioned structure as PROPN as they present the vocative case.

Therefore, in Fig. ~\ref{ex1}, {\tt RT} {\tt @yijialiu}, {\tt \#ACL2017}, {\tt http://url} and {\tt th} are all tagged as X and {\tt :)} is tagged as SYM. 
{\tt rn} is tagged as ADV and {\tt ima} is tagged as PROPN.
We should note that multiple tokens are actually included in one non-explicit contraction, and we use one of their tags as the contraction tag according to their dependency relation or semantic priority.
If there is a hierarchy in the contraction, we use the tag of the head word as the contraction tag. ``Right'' is the dependent of ``now '' in the dependency of {\tt rn}, so we tag {\tt rn} as ADV. Although there is no head word in {\tt im}, we think that the subject is usually more informative than the tense in many downstream tasks such as semantic role labelling, so we use the tag of ``I'', e.g. PROPN as the tag of {\tt im}.  
In Fig. ~\ref{ex2}, {\tt RT} and {\tt $\heart$} would be tagged as VERB, {\tt http://url} as X, {\tt \#Awesome} as ADJ, {\tt \#NAACL}, {\tt @zypandora} and {\tt @nlpnoah} as PROPN. One exception is that we tag URLs always as X regardless whether they are syntactic tokens. The reason is that we observed, in UD\_English, that the URLs are all tagged as X and we want to conform with the UD dataset, although we think it is not reasonable and should be changed to tags such as PROPN when they are treated as syntactic tokens.

\newcite{Gimpel:2011:PTT:2002736.2002747} proposed a twitter POS tagset that handles most of the token level phenomena. However, we found that their methods did not conform with UD conventions and our strategies have some advantages over their tagset. We discuss the main difference as follows.
The first difference lies in the tokenizations.
Gimpel et al. opted not to split contractions or possessives and introduced four new tags (S, Z, L, M) for combined forms: \{nominal, proper noun\} $\times$ \{verb, possessive\}. 
Major concern of designing such tags is to minimize the effort of tokenization, but it is not comprehensive and far from including all of the possible combinations of POS tags within the contractions. 
What's more, only a small proportion of tokens can be categorized into these tags (2.7 \% in total), which casts a doubt of the usefulness of these tags.
Instead UD conventions suggest that we put such complexity into the tokenization module and split contractions and possessives, then tag them accordingly. Therefore, we do not need extra POS tags for the combined forms.
The second difference is that although the phrasal abbreviations are not split in both methods, they use tag G for all the other abbreviations that cannot be tokenized, while we use the tag of the head word or the highest hierarchical word in terms of dependencies as the tag of the whole phrasal abbreviations. We think it is not reasonable to treat the phrasal abbreviations in the same part of speech because obviously abbreviations can have different syntactic functions and we want to preserver the most useful information for both parser and the downstream tasks. For example, {\tt wtf} could be tagged as PRON or INTJ according to the context, and {\tt rn} is usually tagged as ADV. Tagging them merely as G will definitely lose much information.
Third, special tags were designed to handle twitter or online special tokens such as URLs, hashtags and emoticon in \newcite{Gimpel:2011:PTT:2002736.2002747}. However, in most cases, as long as theses tokens are non syntactic, they play the same role in tweets hence should be tagged the same. In this paper, except for the emoticon, tagged as SYM, we tag all the other non-syntactic tokens as X.

@@

\begin{itemize}
\item If the sentence x belongs to has a clear predicate with syntactic function, x should not have any dependent and are attached to this sentence’s predicate;
\item If the sentence x belongs to consists of only non-syntactic tokens we attach x to the first tokens of the sentence;
\item In parataxis case, it might be difficult to decide which sentence or phrase x should belong to. In this case, we always attach x to the previous predicate of the paratactic clause or phrase;
\item x is mostly labelled as \it{discourse}, but URLs are always labelled as \it{list};
\end{itemize}

@@

For syntactic tokens, they are considered the same as normal tokens in the standard text and we annotate them by just following the UD guidelines.

@@

\begin{itemize}
\item Retweet structure: we attach {\tt @user} to {\tt RT} and {\tt RT} to the predicate of the following sentence with {\it discourse} as labels.
\item At-mentioned structure: we attach at-mention username to the predicate of the most relevant sentence with {\it vocative} label. If it is hard to decide which sentence would be the most relevant, we attach it to the previous sentence's predicate.
\item Paratactic parts without punctuations: we follow the UD convention, and set the first part as the main part of the sentence, and set its predicate as the main predicate of the sentence. Every predicate of the following parts should attach to the main predicate. If the main part and some following part are both phrases, the label is {\it list}, otherwise the label is {\it parataxis}. Note that we still have an exception for URLs, conforming with UD, which is URLs are always labelled as {\it list}, regardless of the type of the other part.
\end{itemize}

% unlabelled, Informal
% exclude non-syntactic tokens, punctuations
% Annotation conventions
\newcite{kong-EtAl:2014:EMNLP2014} built a syntactic treebank of tweets called Tweebank using Gimpel et al.'s POS tagset. 
Apart from the difference in POS tagging, we found the following difference in both methods.
First, Tweebank is an unlabelled dependency treebank, while we are annotating labelled dependencies. There is clearly a doubt about the usefulness of the Tweebank. We annotate the labelled dependencies by following the UD annotation guideline.
Second, most of Tweebank was built in a day by two dozen annotators, most of whom had only cursory training in the annotation scheme~\cite{kong-EtAl:2014:EMNLP2014}. We think, after investigation, that despite the speed, the quality of data is still to be improved. We follow strictly the UD annotation guideline and only make twitter specific conventions when there is no suitable convention to adopt in UD guideline or no relevant data found in UD\_English.
Third, \newcite{kong-EtAl:2014:EMNLP2014} developed a first-order sequence model to filter out the non-syntactic tokens and punctuations, and exclude them before parsing.
It is very different from most of the annotation conventions, especially for UD conventions. We believe that every token should be included in the dependency tree, whether they have syntactic functions or are to be evaluated in the end, as in most of the conventions, so we annotate all tokens.
Last, \newcite{kong-EtAl:2014:EMNLP2014} adopt the ``Yamada-Matsumoto" conventions~\cite{Yamada03statisticaldependency}, where auxiliary verbs are parents of main verbs, and prepositions are parents of their arguments, in contrast to UD conventions, where content words are primary tokens, and dependency relations hold primarily between them, rather than being indirect relations mediated by function words.

>>>>

\subsection{Linguistic Phenomena of Twitter} \label{introduction}
%Eisenstan
As the representative of the web language, different from the standard languages such as newswire, twitter has its own linguistic phenomena, that could be categorized into token level and structure level phenomena, and we will discuss them separately.

@@

There are special tokens that will appear only or much more frequently in tweets. According to the syntactic functions these tokens have in the tweets, we classify them into three types: syntactic tokens, non-syntactic tokens and optionally syntactic tokens. 
Syntactic tokens are just like normal tokens in the standard text. They always have syntactic functions in tweets. They could be orthographic variants or abbreviations that represents a phrase.
Non-syntactic tokens are tokens that will not have syntactic function in tweets, and they are not considered in the syntax analysis and need to be treated independently.
Optionally syntactic tokens are tokens that can have either syntactic or non-syntactic functions based on the context. Usually they are tokens like URLs ad hashtags that do not participate in the syntax of the tweets, and we treat them as non-syntactic functions by default. However, similar in the analysis of \newcite{Gimpel:2011:PTT:2002736.2002747} and \newcite{kong-EtAl:2014:EMNLP2014}, in some cases they are perceived as the arguments or predicates of tweets and therefore have syntactic functions. In this case we treat them as syntactic tokens.

@@

\begin{itemize}
\item syntactic tokens
	\begin{itemize}
		\item phrasal abbreviations: {\tt wtf} (what the fuck), {\tt rn} (right now), etc.
		\item orthographic variants: {\tt u} (you), {\tt sooooo}, {\tt dat} (that), etc.
	\end{itemize}
\item Non-syntactic tokens
	\begin{itemize}
		\item truncated tokens: last tokens that are truncated in tweets exceeding twitter character limits. Although It might be possible to recover the original tokens from the context in some cases, it is still extremely hard to further predict the rest of the truncated part.
	\end{itemize}
\item Optionally syntactic tokens
	\begin{itemize}
		\item retweet discourse marker: RT
		\item URL: http://bit.ly/xyz
		\item hashtag: \#ACL
		\item at-mentioned username: @user
		\item emoticons and emojis: :), :-), $\heart$, etc.
	\end{itemize}
\end{itemize}

As optionally syntactic tokens will be treated as syntactic or non-syntactic tokens based on the specific context, for the sake of simplicity, if not necessary, we will only discuss the syntactic and non-syntactic tokens in the following sections.

@@

\begin{itemize}
\item Retweet structure: {\tt RT @user : \{tweet content\}} is a typical structural pattern when a user is retweeting from other users. 
\item At-mentioned structure: Except the retweet structure and syntactically at-mentioned username, we think that all of other at-mentions are vocative case, usually appearing at the beginning or the end of the sentences in tweets.
\item Paratactic parts without punctuations: We use ``part'' to refer to a self-contained clause or phrase, which is independent of other ``part''. Very often, one tweet is comprised of multiple parts without any delimiting punctuations, such as {\tt \{part 1\} \{part 2\} ...}
\end{itemize}

We treat retweet structure and at-mentioned structure as structural patterns and keep their annotations consistent across tweets. For parataxis case, we treat each single sentence or phrase normally, and connect them together following the UD conventions.

@@

From Fig.~\ref{ex1}, we argue that {\tt u} and {\tt rn} are syntactic tokens.
{\tt th} is a truncated token and does not have any syntactic function, although we can infer that the {\tt th} could be probably {\tt the}.
{\tt \#ACL2017}, {\tt :)} and {\tt http://url} are optionally syntactic tokens, and should be all treated as non-syntactic tokens in this example.
{\tt RT @Yijia :} should be analyzed as the retweet structure, and {\tt heading to Canada for ACL} and {\tt can u gimme some money} should be considered as paratactic clauses without punctuations.

@@

\subsection{POS Tagging}
We follow the UD V2 morphology guideline and use the Universal POS tags~\cite{PETROV12.274} to tag the tweet tokens. For twitter specific linguistic phenomena, we have set different strategies and try to align with the UD conventions and UD\textunderscore English as close as possible.\footnote{\url{https://github.com/UniversalDependencies/UD\_English}}

@@

In structure level phenomena, we think {\tt RT @user} in retweet structure have no contribution to the tweet syntax and tag both of them as X.
We tag usernames in at-mentioned structure as PROPN as they present the vocative case.

In Fig.~\ref{ex1}, {\tt RT} {\tt @yijialiu}, {\tt \#ACL2017}, {\tt http://url} and {\tt th} are all tagged as X and {\tt :)} is tagged as SYM. 
{\tt rn} is tagged as ADV.
We should note that multiple tokens are actually included in a phrasal abbreviation, and we use one of their tags as the tag of the whole abbreviation according to their dependency relation or semantic priority.
If there is a hierarchy in the abbreviation, we use the tag of the head word. In UD, ``Right'' is the dependent of ``now '' in the dependency of {\tt rn}, so we tag {\tt rn} as ADV. 
%Although there is no head word in {\tt im}, we think that the subject is usually more informative than the tense in many downstream tasks such as semantic role labelling, so we use the tag of ``I'', e.g. PROPN as the tag of {\tt im}.  
In Fig.~\ref{ex2}, {\tt RT} and {\tt $\heart$} would be tagged as VERB, {\tt http://url} as X, {\tt \#Awesome} as ADJ, {\tt \#NAACL}, {\tt @zypandora} and {\tt @nlpnoah} as PROPN. One exception is that we tag URLs always as X regardless whether they are syntactic tokens. The reason is that we observed, in UD\_English, that the URLs are all tagged as X and we want to conform with the UD dataset, although we think it is not appropriate and should be changed to tags such as PROPN when they are treated as syntactic tokens.

\newcite{Gimpel:2011:PTT:2002736.2002747} proposed a twitter POS tagset that handles most of the token level phenomena. However, we found that their methods did not conform with UD conventions and our strategies have some advantages over their tagset. We discuss the main difference as follows.
The first difference lies in the tokenizations.
\newcite{Gimpel:2011:PTT:2002736.2002747} opted not to split contractions or possessives and introduced four new tags (S, Z, L, M) for combined forms: \{nominal, proper noun\} $\times$ \{verb, possessive\}. 
Major concern of designing such tags is to minimize the effort of tokenization, but it is not comprehensive and far from including all of the possible combinations of POS tags within the contractions. 
What's more, only a small proportion of tokens can be categorized into these tags (2.7 \% in total), which casts a doubt of the usefulness of these tags.
Instead UD conventions suggest that we put such complexity into the tokenization module and split contractions and possessives, then tag them accordingly, so that we do not need extra POS tags for the combined forms.
The second difference is that although the phrasal abbreviations are not split in both methods, they use tag G for all the phrasal abbreviations, while we use the tag of the head word or the highest hierarchical word in terms of dependencies as the tag of the whole phrasal abbreviations. 
We think it is not reasonable to treat the phrasal abbreviations in the same part of speech because obviously abbreviations can have different syntactic functions and we want to preserve the most useful information for both parser and the downstream tasks. For example, {\tt wtf} could be tagged as PRON or INTJ according to the context, and {\tt rn} is usually tagged as ADV. Tagging them merely as G will definitely lose much information.
Third, special tags were designed to handle twitter or online special tokens such as URLs, hashtags and emoticon in \newcite{Gimpel:2011:PTT:2002736.2002747}. However, in most cases, as long as theses tokens are non syntactic, they play the same role in tweets, hence should be tagged the same. In this paper, except for the emoticon and emoji, tagged as SYM, we tag all the other non-syntactic tokens as X.

@@

\begin{itemize}
\item If the sentence x belongs to has a clear predicate with syntactic function, x should not have any dependent and is attached to this sentence’s predicate;
\item If the sentence x belongs to consists of only non-syntactic tokens we attach x to the first token of the sentence;
\item In parataxis case, when it is difficult to decide which sentence or phrase x should belong to, we always attach x to the previous predicate of the paratactic clause or phrase;
\item x is mostly labelled as {\it discourse}, but URLs are always labelled as {\it list};
\end{itemize}

@@

For syntactic tokens, they are considered the same as normal tokens and we just annotate them by following the UD guidelines.

@@

\begin{itemize}
\item Retweet structure: we attach {\tt @user} to {\tt RT} and {\tt RT} to the predicate of the following sentence with {\it discourse} as labels.
\item At-mentioned structure: we attach at-mentioned username to the predicate of the most relevant sentence with {\it vocative} label. If it is hard to decide which sentence would be the most relevant, we attach it to the previous sentence's predicate.
\item Paratactic parts without punctuations: we follow the UD convention, set the first part as the main part of the sentence, and set its predicate as the main predicate of the sentence. Every predicate of the following parts should attach to the main predicate. If the main part and some following part are both phrases, the label is {\it list}, otherwise the label is {\it parataxis}. Note that we still have an exception for URLs, conforming with UD, which are always labelled as {\it list}, regardless of the type of the other part.
\end{itemize}

% unlabelled, Informal
% exclude non-syntactic tokens, punctuations
% Annotation conventions
\newcite{kong-EtAl:2014:EMNLP2014} built a syntactic treebank of tweets called Tweebank using Gimpel et al.'s POS tagset. 
Apart from the difference in POS tagging, we found the following difference in both methods.
First, Tweebank is an unlabelled dependency treebank. There is clearly a doubt about the usefulness of the Tweebank. We annotate the labelled dependencies by following the UD annotation guideline.
Second, most of Tweebank was built in a day by two dozen annotators, most of whom had only cursory training in the annotation scheme~\cite{kong-EtAl:2014:EMNLP2014}. We think, after investigation, that despite the speed, the quality of data is still to be improved. We follow strictly the UD annotation guideline and only make twitter specific conventions when there is no suitable convention to adopt in UD guideline or no relevant data found in UD\_English.
Third, \newcite{kong-EtAl:2014:EMNLP2014} developed a first-order sequence model to filter out the non-syntactic tokens and punctuations, and exclude them before parsing.
It is very different from most of the annotation conventions, especially for UD conventions. We believe that every token should be included in the dependency tree, whether they have syntactic functions or are to be evaluated in the end, as in most of the conventions, so we annotate all tokens.
Last, \newcite{kong-EtAl:2014:EMNLP2014} adopt the ``Yamada-Matsumoto" conventions~\cite{Yamada03statisticaldependency}, where auxiliary verbs are parents of main verbs, and prepositions are parents of their arguments, in contrast to UD conventions, where content words are put into the primacy, and dependency relations hold primarily between them, rather than being indirect relations mediated by function words. Auxiliary verbs and prepositions are all function words, and are attached to the most related content words in UD.
